---
title: "Accessing model-output data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

An important function of `hubUtils` is allowing for the connection to data in the `model-output` directory to facilitate interrogation, filtering and extraction of Hub data.

There are two functions for connecting to `model-output` data:

- `connect_hub()` is used for connecting to fully configured hubs (i.e. which contain valid `admin.json` and `tasks.json` in a `hub-config` directory). This function uses configurations defined in config files in the `hub-config/` directory and allows for connecting to hubs with files in multiple file formats (allowable formats specified by the `file_format` property of `admin.json`).
- `connect_model_output()` allows for connecting directly to the contents of a `model-output` directory and is useful for connecting to appropriately organised files in an informal hub (i.e. which has not been fully configured with appropriate `hub-config/` files.)

Both functions establish connections through the [`arrow`](https://arrow.apache.org/docs/r/) package, specifically by opening datasets as [`FileSystemDataset`s](https://arrow.apache.org/docs/r/reference/Dataset.html), one for each file format. 

Where multiple file formats are accepted in a single Hub, file format specific `FileSystemDataset`s are combined into a single `UnionDataset` for single point access to the entire Hub `model-output` dataset. 
This only applies to `connect_hub()` in fully configured Hubs, where config files can be used to determine a unifying schema across all file formats. 

In contract, `connect_model_output()` can only be used to open single file format datasets of the format defined explicitly through the `file_format` argument.


```{r setup}
library(hubUtils)
library(dplyr)
```

## Connecting to a configured hub


```{r}
hub_path <- system.file("testhubs/flusight", package = "hubUtils")
hub_con <- connect_hub(hub_path)
hub_con
```

To access data from a hub connection you can use dplyr verbs and construct querying pipelines.

```{r}
hub_con %>%
    filter(output_type == "quantile", location == "US") %>%
    collect()
```

Note however that not all dplyr filtering options are available for all data types yet.

For example, if you wanted to get all quantile predictions for the last forecast date in the hub, you might try:

```{r, error=TRUE}
hub_con %>%
    filter(output_type == "quantile", location == "US") %>%
    filter(forecast_date == max(forecast_date)) %>%
    collect()
```

This doesn't work however as `arrow` does not have an equivalent `max` method for `Date[32]` data types.

In such a situation, you could collect after applying the first filtering level which does work for arrow and then finish the filtering on the in-memory data returned by collect.

```{r}
hub_con %>%
    filter(output_type == "quantile", location == "US") %>%
    collect() %>%
    filter(forecast_date == max(forecast_date)) 
```

Alternatively, depending on the size of the data, in might be quicker to filter the data in two steps:

1. get the last forecast date available for the filtered subset.
2. use the last forecast date in the filtering query.

```{r}
last_forecast <- hub_con %>%
    filter(output_type == "quantile", location == "US") %>%
  pull(forecast_date, as_vector = TRUE) %>%
  max()


hub_con %>%
    filter(output_type == "quantile", location == "US",
           forecast_date == last_forecast) %>%
    collect()
```



## Connecting directly to a `model_output_dir`

There is also an option to connect directly to a model output directory without using any metadata in a hub config file. This can be useful when a hub has not been fully configured yet. 

The approach does have certain limitations though. For example, an overall unifying schema cannot be determined from the config files so the ability of `open_dataset()` to connect and parse data correctly cannot be guaranteed across files.

In addition, only a single file_format dataset can be opened.

```{r}
model_output_dir <- system.file("testhubs/simple/model-output", package = "hubUtils")
mod_out_con <- connect_model_output(model_output_dir, file_format = "csv")
mod_out_con
```


```{r}
mod_out_con %>%
    filter(output_type == "quantile", location == "US") %>%
    collect()
```

