---
title: "Setting up a Hub"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r setup}
library(hubUtils)
```


## Create Hub Repository from Template

The first step to setting up a Modeling Hub is to make a copy of the [`hubTemplate`](https://github.com/Infectious-Disease-Modeling-Hubs/hubTemplate) GitHub template repository and clone a local copy of it.


## Configure Modeling Hub

Hubs are configured through two JSON config files which are used to validate new submissions as well as inform data users of available data.
- `admin.json`: Contains Hub wide administrative configuration settings
- `tasks.json`: Contains round-specific metadata of modeling tasks including task, target and output type metadata as well as details of submission windows.
These files should live in a directory called `hub-config/` in the root of the Hub repository.


For more details on these files, guidance on how to set them up and access to templates, please see our central [**hubDocs**](https://github.com/Infectious-Disease-Modeling-Hubs/hubDocs) as well the [JSON schema](https://github.com/Infectious-Disease-Modeling-Hubs/schemas) the config files should adhere to.


## Validate Config files

You can use function `validate_config()` to check whether your Hub config files are valid. To specify the file you want to validate you can either provide the path to the root of the Hub to argument `hub_path` (which assumes the config files are correctly located in directory `hub-config/`) or you can provide a direct path to a config file to argument `config_path`. You also need to specify the type of config file through argument `config` (one of `"tasks"` or `"admin"`, defaults to `"tasks"`).

The function will validate a given config file against a specific version of it's schema, specified through argument `schema_version`. The default value of `schema_version` is `"from_config"` which uses the version specified in the `schema_version` property of the config file being validated. 


```{r}
validate_config(
    hub_path = system.file("testhubs/simple/", package = "hubUtils"),
    config = "tasks"
)
```

If validation succeeds, the function returns `TRUE`. The path to the config file validated and the version and URL of the schema used for validation are also attached as attributes `"config_path"`, `"schema_version"` and `"schema_url"` respectively.


You can validate a config file against the latest version of the schema by using `"latest"` or you can choose a specific version, e.g `"v0.0.1"`. The fuunction defaults to using stable schema versions released to the `main` branch, but you can choose to validate against another branch (e.g. an upcoming development version).

```{r}
validate_config(
    hub_path = system.file("testhubs/simple/", package = "hubUtils"),
    config = "tasks",
    schema_version = "v0.0.0.9"
)
```


```{r}
validate_config(
    hub_path = system.file("testhubs/simple/", package = "hubUtils"),
    config = "tasks",
    schema_version = "latest"
)
```


### Validation returning errors

If validation of the config file fails, the function returns `FALSE`. An additional list dataframe of errors returned by the `ajv` validation engine used is also attached as attribute `"errors"`.


```{r}
config_path <- system.file("error-schema/tasks-errors.json", 
                           package = "hubUtils")
```

```{r, eval=FALSE}
validate_config(config_path = config_path, config = "tasks")
```


```{r, echo=FALSE}
out <- validate_config(config_path = config_path, config = "tasks", pretty_errors = FALSE)
out
```


Because the default output of the validator can be unwieldy and difficult to review, the function also prints a more user friendly and concise version of the errors table. 


```{r, echo=FALSE}
hubUtils:::launch_pretty_errors_report(out) 
```

In the example above: 

- `instancePath` indicates the location of the validation error in the config file.
- `schemaPath` indicates the location of the element in the schema which is failing validation.
- `keyword` indicates the keyword causing the validation error.
- `message` is the validation error message returned by the validator.
- `schema` describes the valid schema values the failing keyword should conform to.
- `data` is the value of the property in the config file which is failing validation.

You can suppress printing the pretty errors table with `pretty_errors = FALSE`.

```{r, eval=FALSE}
validate_config(config_path = config_path, config = "tasks",
                pretty_errors = FALSE)
```

